---
title       : Predicting Gas Mileage in Cars
subtitle    : An App to estimate your fuel economy
author      : Mohar Guha
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Problem Statement

* With the recent surge in gas prices we would like to identify the features in a car that affect gas mileage. 

* Consider the following features of the car for building this app.

  * Weight, Horsepower (numeric) 
  * Transmission, Number of Cylinders, Number of Gears (categorical)

## Methodology

* Load the dataset 'mtcars' for analysis (type ?mtcars in R for documentation)

* Exploratory analysis - quick graphs to test our claim.

* Fit regression model

* Use the model for prediction

--- .class #id 

 
## Exploratory analysis 

```{r out.width = '50%', echo = F, message = F} 
library(lattice)
library(ggplot2)
p0<-xyplot(mpg ~ wt | factor(cyl), data=mtcars, pch=16,layout=c(3,1),groups = factor(am,labels=c("Auto","Manual")),type=c("p","g"),xlab='Weight (1000 lb)', ylab='Miles per Gallon', main="MPG vs.Weight by Cylinder and Transmission",auto.key = TRUE)
mtcars$amf[mtcars$am == 0] <- "automatic"
mtcars$amf[mtcars$am == 1] <- "manual"
p1<-qplot(x = hp, y = mpg, facets = . ~ amf, xlab = "Horsepower", ylab = "Miles 
          Per Gallon", data = mtcars, main = "MPG vs.Horsepower by Transmission")
p0
p1
``` 

---

## Regression Model
$\rm mpg=\beta_0 \rm weight + \beta_1 \rm transmission+\beta_3 \rm cylinders +
\beta_4\rm gear+\beta_5 \rm horsepower +\rm error$
```{r echo=F,results='asis'}
library(xtable)
model1<- lm(mpg~wt+as.factor(am)+as.factor(cyl)+as.factor(gear)+hp,data=mtcars)
print(xtable(model1), type = "html")
```

---
## Results
* Mileage is directly proportional to weight
* Manual transmission is fuel efficient
* Inverse relationship between horse power and mpg
* A 4 cylinder car and a manual transmission then you get the best gas mileage.


Have fun playing with this app!








